---
title: 'Assignment 2'
author: "Your Name and Surname"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Installing necessary package(s)
install.packages(...)
library(...)
```


# Dataset

The `BreastCancer` dataset contains cytological measurements of breast tissue 
samples, used to classify tumors as benign or malignant. It includes numerical 
features describing cell characteristics derived from microscopic examination.

To load the dataset to your environment, just run the code below:

```{r message=FALSE, warning=FALSE}
install.packages("mlbench")   
library(mlbench)
data("BreastCancer")          
BreastCancer$Cl.thickness <- as.numeric(BreastCancer$Cl.thickness)
```


# Drawing a plot for distribution

1. Please draw a plot to visualize the distribution 
of tissue thickness (`Cl.thickness`) of the patients according to `Class` 
indicating whether the tumor in the breast tissue is benign or malignant.
(40 pts).

```{r}

```


2. Please solve the problems in the plot and make it better by adding 
some information (30 pts).

```{r}

```


3. Interpret the plot (30 pts). 

...
