---
title: 'Assignment 1'
author: "Your Name and Surname"
date: "2025-10-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Installing necessary package(s)
install.packages(...)
library(...)
```

# Dataset

The dataset `eskisehir` contains the total population for the top-5 crowded
districts of Eskisehir in 2024.

```{r}
eskisehir <- data.frame(
  district   = c("Odunpazari", "Tepebasi", "Sivrihisar", "Cifteler", "Seyitgazi"),
  population = c(426581, 394734, 20258, 14814, 12878))
```

# Drawing a barplot

1. Please draw a boxplot to visualize the population of districts (20 pts).

```{r}
ggplot(...) + 
  geom_bar(...)
```

2. Please solve the problems in the plot to make it better (20 pts).

```{r}

```

3. Then configure the plot to present it in a better way (40 pts).
You can use the tricks that you learned in the lecture.

```{r}

```

4. Interpret the plot (20 pts). It must be about the information of the 
district populations not the technical part of the plot.

...

