---
title: "Exercises on Sep 29, 2025"
author: "Mustafa Cavus"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

R Markdown is a powerful tool for combining **text, code, and results** in a single document.  
For data analysts and data scientists, it offers several key advantages:

- **Reproducibility:** Your analysis and results are generated directly from the code, ensuring they can be reproduced.  
- **Documentation:** You can explain your workflow, methods, and insights alongside the code.  
- **Communication:** Easily share reports with colleagues, managers, or the public in PDF, HTML, or Word formats.  
- **Efficiency:** Inline R code allows you to dynamically display calculations, summaries, and results within your text.


# 2. Markdown Basics

You can use Markdown to add text and structure to your document.

## 2.1. Headings

Use # for headings.

## 2.2. Emphasis

italic with `*text*` → *text*

bold with `**text**` → **text**

## 2.3. Lists

* Item 1
* Item 2
* Item 3

## 2.4. Links and Images

`[R Markdown Guide](https://rmarkdown.rstudio.com/)` → [R Markdown Guide](https://rmarkdown.rstudio.com/)


## 2.5. Code Chunks

You can insert R code inside triple backticks with `{r}`:

```{r echo=FALSE}
x <- c(1, 2, 3, 4, 5)
mean(x)
```

backticks:

* Windows: AltGr + ‘ (apostrophe key, to the left of Enter)

* Mac: Option + ’ (apostrophe key)

In RStudio, you can also insert a new code chunk by clicking the +c button at the right top of the editor.

You can control how chunks behave using options inside the chunk. For example:

```{r}
# Code runs, but is not shown
mean(1:10)
```

* `echo=FALSE` → hides the code, but shows the results.
* `message=FALSE`, `warning=FALSE` → suppress messages and warnings.
* or use the shortcut "gear icon" on the right top of the chunk.

# Your Turn (5 mins)
1. Insert a new code chunk using the RStudio shortcut.
2. In that chunk, run summary(mtcars) but hide the code.


## 2.7. Inline Codes

If you want to display the mean of the vector `x <- c(1,2,3,4,5)` directly in your text, you can use inline

The mean of the vector `x <- c(1,2,3,4,5)` is r mean(c(1,2,3,4,5)) within backticks.


# 3. Data Visualization with {ggplot2}

The **ggplot2** package is used to create graphics in R.  
It follows the idea of building plots layer by layer.  

```{r include=FALSE}
install.packages("ggplot2") # install package
library(ggplot2)            # load package
```

## 3.1. Basic Structure

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```


# Your Turn  (10 mins)
1. Create a boxplot of the variable `mpg` from the `mtcars` dataset.
2. Describe the insights you can get from the plot.


## 3.2. Customizing Plots

After creating a basic plot, you can customize it in several ways:  

### 3.2.1. Adding a Title and Axis Labels

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Car Weight vs. MPG",
       x     = "Weight (1000 lbs)",
       y     = "Miles per Gallon")
```


### 3.2.2. Changing the Theme

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  theme_minimal()   # alternatives: theme_gray(), theme_classic(), theme_light()
```

* `theme_minimal()` → clean background
* `theme_classic()` → white background with axes lines
* `theme_light()` → light gray background


### 3.2.3. Changing the Theme

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl), shape = factor(gear))) +
  geom_point(size = 3) +
  labs(color = "Cylinders", 
       shape = "Gears")
```

* `color` → color points by a variable
* `shape` → change point shape based on a variable
* `size` → adjust point size

# Your Turn  (15 mins)
1. Create a scatter plot of `hp` vs `mpg`.
2. Add a plot title and custom x/y labels.
3. Change the background using your favorite one.
4. Color the points by `cyl`.