---
title: "Visualizing Amounts"
author: "Mustafa Cavus"
date: "2025-10-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE} 
# You can hide the messages and warnings when installing a package.
# This is important when installing a package to avoid unnecessary installation 
# messages and ensure that the document contains only the necessary information.

install.packages("ggplot2") # to use ggplot visualization tools
install.packages("dplyr") # to use the new pipe (|>) operator and perform data 
# manipulation operations
install.packages("tidyverse")
library(ggplot2)
library(dplyr)
library(tidyverse)
```


# How to draw a barplot

Let's create a data abject called `movie`. It consists two variables `name` and 
`revenue`. The variables are nominal and numeric, respectively. It means we can
visualize the revenues of these movies by using a barplot.

```{r}
movie <- data.frame(name = c("Star Wars",
                             "Jumanji",
                             "Pitch Perfect 3",
                             "Greatest Showman",
                             "Ferdinand"),
                    revenue = c(71565498,
                                36169328,
                                19928525,
                                8805843,
                                7316746))
```

To draw a barplot, we need to create two layers: the first one indicates the 
data, and the variables that we want to put to the axes. Then we must specify
the geom layer which shows the plot type: `geom_bar()`.

```{r eval=FALSE, include=FALSE}
ggplot(movie, aes(x = name, y = revenue)) +
  geom_bar()
```

The code above returns an error because we did not specify the stat argument 
in the `geom_bar()` layer. The stat argument is typically used with the options 
`identity` or `count`.

```{r}
ggplot(movie, aes(x = name, y = revenue)) +
  geom_bar(stat = "identity")
```

In the plot above, you will notice that the values on the y-axis are in a 
different format. For example, 2e+07 is converted to decimal form as 2 x 10^7. 
Converting all values on the axis to decimal form allows us to obtain a more 
readable plot.

```{r}
ggplot(movie, aes(x = name, y = revenue)) +
  geom_bar(stat = "identity") + 
  scale_y_continuous(labels = scales::comma)
```

Another step we need to take for this plot to be more easily readable is to 
sort the bars in  ascending or descending order based on their lengths. We can 
decide which option to choose according to the content of the message.

```{r}
ggplot(movie, aes(x = reorder(name, -revenue), y = revenue)) +
  geom_bar(stat = "identity") + 
  scale_y_continuous(labels = scales::comma)
```

You will see that the `reorder()` function we use to sort the bars causes an 
error message for the solution we use to change the format of the label values 
on the y-axis. To solve this problem, you can use the solution below:

```{r}
ggplot(movie, aes(x = reorder(name, +revenue), y = revenue)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

As you can see, there can be multiple solutions to a problem that arises, and 
in some cases, not all solutions may work. When you encounter such issues, 
remember that your first resource should be Google or stackoverflow.com.

The last step we need to take for this plot is to arrange the axis titles and 
add the plot title.

```{r}
ggplot(movie, aes(x = reorder(name, +revenue), y = revenue)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) + 
  labs(x       = "Movie",
       y       = "Revenue ($)",
       title   = "Highest Grossing Movies from December 22 to 24, 2017",
       caption = "Data source: Box Office Mojo") +
  theme_minimal()
```

One of the main issues you may encounter in such plots is that the bar labels 
on the axis containing the categorical variable can be very long, leading to 
overlapping and making them difficult to read.A common but poor practice to 
solve this problem is to rotate the bar labels by 45 or 90 degrees, 
which actually makes them harder to read.

```{r}
ggplot(movie, aes(x = reorder(name, +revenue), y = revenue)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) + 
  labs(x       = "Movie",
       y       = "Revenue ($)",
       title   = "Highest Grossing Movies from December 22 to 24, 2017",
       caption = "Data source: Box Office Mojo") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```

Instead, switching the positions of the axes is a better solution to avoid 
making the plot harder to read.

```{r}
ggplot(movie, aes(x = reorder(name, +revenue), y = revenue)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) + 
  labs(x       = "",
       y       = "Revenue ($)",
       title   = "Highest Grossing Movies from December 22 to 24, 2017",
       caption = "Data source: Box Office Mojo") + 
  coord_flip() +
  theme(text = element_text(size = 20))
```


# How to draw a grouped barplot

Visualization of the `US_median_income` dataset found on page 49 of Wilke's 
book. The book’s source code and the dataset is hosted on GitHub, 
https://github.com/clauswilke/dviz.supp/tree/master/data.
Please first download the data, and follow the steps below to work on a dataset
you want to call the outside of the environment:

* (If you use local RStudio skip this step) Upload the data file your 
environment using the "upload" button on the files window.
* The file will appear within files, then left click and load the data to 
your environment, and follow the steps below.

```{r}
load("/cloud/project/US_income_age.rda")
```

You need to convert the continuous `age` variable in the `US_income_age` 
dataset into a categorical variable. For this, you can use the `within()` 
function.

```{r}
US_age_cat <- within(US_income_age, {   
  age.cat <- NA 
  age.cat[25 <= age & age <= 34] <- "25-34"
  age.cat[35 <= age & age <= 44] <- "35-44"
  age.cat[45 <= age & age <= 54] <- "45-54"
  age.cat[55 <= age & age <= 64] <- "55-64"
  age.cat[65 <= age & age <= 74] <- "65-74"})

final_US_income_age <- US_age_cat |> 
  drop_na() |> 
  group_by(age.cat) |> 
  summarise(moi = median(income))

ggplot(final_US_income_age, aes(x = age.cat, y = moi)) + 
  geom_bar(stat = "identity")
```

If we try to draw the plot above without using the `drop_na()` function while 
preparing the data, that is, without excluding missing observation values, 
you will see a bar labeled NA on the x-axis.

We need to make a few adjustments to the plot to give it its final form:

```{r}
ggplot(final_US_income_age, aes(x = age.cat, y = moi)) + 
  geom_bar(stat = "identity") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```

If we try to sort the bars in the plot above in ascending or descending order 
based on the values they represent, it becomes difficult to interpret the 
income changes based on age, which we have transformed into an ordered 
categorical variable. Drawing conclusions about income changes based on the 
increase or decrease in age becomes challenging. However, this situation may 
not always hold true. Even though the relevant variable is ordered and 
categorical, we can perform the sorting operation based on the message we want 
to convey, contrary to its structure. I want to warn you that this situation 
is not something we will encounter very frequently in practice.

```{r}
ggplot(final_US_income_age, aes(x = reorder(age.cat, +moi), y = moi)) + 
  geom_bar(stat = "identity") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```

Let's create a grouped bar chart using the same dataset. For this, we need to 
prepare the data again. Let's focus on some states for the grouping. We can 
select the observations corresponding to the states using the `filter()` 
function, and then use `group_by()` to calculate the median income for the 
grouped data by age and state, allowing us to access the data we will use 
for visualization.

```{r}
final_US_income_age2 <- US_age_cat |>
  drop_na() |> 
  filter(state %in% c("Michigan", "Colorado", "Virginia", "Texas")) |>
  group_by(age.cat, state) |> 
  summarise(moi = median(income))
```

To draw a grouped bar chart, we need to use two arguments. In `aes()`, we 
should specify the variable for grouping using the `fill` argument, and in 
the `geom_bar()` layer, we need to set the `position` argument to the value 
`dodge`.

```{r}
ggplot(final_US_income_age2, aes(x = age.cat, y = moi, fill = state)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       fill    = "State",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```

We can sort the age groups in ascending or descending order using the 
`reorder()` function.

```{r}
ggplot(final_US_income_age2, aes(x = reorder(age.cat, +moi), y = moi, fill = state)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       fill    = "State",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```

Alternatively, you can obtain a multi-panel plot based on a desired variable 
by using the `facet_wrap()` layer.

```{r}
ggplot(final_US_income_age2, aes(x = age.cat, y = moi)) + 
  geom_bar(stat = "identity") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       fill    = "State",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  facet_wrap(~final_US_income_age2$state) + 
  theme_bw()
```

Let's create a stacked bar plot using the same dataset. Here, we can use the 
dataset we prepared for the grouped bar plot: The only different step we need 
to take to draw the plot is to specify the 'position' argument in the 
`geom_bar()` layer as "stack".

```{r}
ggplot(final_US_income_age2, aes(x = age.cat, y = moi, fill = state)) + 
  geom_bar(stat = "identity", position = "stack") + 
  labs(x       = "Age",
       y       = "Median Income ($)",
       fill    = "State",
       title   = "2016 Median Income by Age Ranges of American Households",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```


# As an alternative of barplot is dotplot

We will use the `US_income_age` dataset for the applications in this section.
Let's start by drawing a bar plot before the dot plot.

```{r}
final_US_income_age3 <- US_income_age |>
  drop_na() |>
  group_by(state) |>
  summarise(mean_income = mean(income))

ggplot(final_US_income_age3, aes(x = mean_income, y = state)) + 
  geom_bar(stat = "identity") + 
  labs(x = "Mean Income ($)",
       y = "State",
       fill = "State",
       title = "Average Income of American Households by State in 2016",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()

```

Since it is difficult to read the plot, the bars need to be sorted according 
to the values they represent.

```{r}
ggplot(final_US_income_age3, aes(x = mean_income, y = reorder(state, +mean_income))) + 
  geom_bar(stat = "identity") + 
  labs(x = "Median Income ($)",
       y = "State",
       fill = "State",
       title = "Average Income of American Households by State in 2016",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```

Despite making all the necessary adjustments we have seen so far on this plot, 
the close values of the states and the tightly packed long bars do not create 
a good (distinguishable) appearance. In such cases, using a dot plot instead 
of a bar plot is a better choice. We can use the geom_point() layer to create 
the dot plot.

```{r}
ggplot(final_US_income_age3, aes(x = mean_income, y = reorder(state, +mean_income))) + 
  geom_point() + 
  labs(x = "Median Income ($)",
       y = "State",
       fill = "State",
       title = "Average Income of American Households by State in 2016",
       caption = "Data source: US Census Bureau (https://github.com/clauswilke/dviz.supp/blob/master/data/US_income_age.rda)") + 
  theme_bw()
```


