---
title: 'Assignment 3'
author: "Your Name and Surname"
date: "2025-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Installing necessary package(s)

```


# Dataset

The `salary` dataset contains information about the average annual salaries of 
data professionals across different roles and experience levels. It includes 
numerical data on salaries (`mean_salary`) in Rupees as well as 
categorical variables describing the job role (`position`) and 
experience level (`experience`). The experience levels are ordered from 
Junior → Intermediate → Senior → Executive, reflecting increasing levels of 
expertise and responsibility. This dataset can be used to explore which 
positions and experience levels contribute most to overall salary distribution 
in the data science field.

Source: [Kaggle – Data Science Fields Salary Categorization](https://www.kaggle.com/datasets/whenamancodes/data-science-fields-salary-categorization?resource=download) (accessed October 2025).

To load the dataset into your environment, just run the code below:

```{r}
salary <- data.frame(
  position    = c("Data Analyst", "Data Analyst", "Data Analyst", "Data Analyst",
                  "Data Engineer", "Data Engineer", "Data Engineer", "Data Engineer",
                  "Data Scientist", "Data Scientist", "Data Scientist"),
  experience  = c("Junior", "Executive", "Intermediate", "Senior",
                  "Junior", "Executive", "Intermediate", "Senior",
                  "Junior", "Intermediate", "Senior"),
  mean_salary = c(4293623, 9548340, 5705070, 8905628,
                  4689309, 19534312, 6841836, 10903873,
                  4402653, 6527813, 12171827))
```


# Drawing a plot for proportion

1. Draw a plot showing the proportion of total mean salaries by experience and 
position. The plot should clearly display the hierarchical contribution 
of each job title within its experience level.

```{r}

```


2. Interpret the plot (30 pts). 


